WIP: Transparent Standby for Low-Power, Resource-Constrained Embedded Systems

### 2018-04-12 LCTES/body.tex
(>) Revisões artigo LCTES

\CEU is a Esterel-based~\cite{ceu.tecs17} reactive programming language
targeting resource-constrained embedded systems~\cite{ceu.sensys13}.
%
It is grounded on the synchronous concurrency model, which has been
successfully adopted in the context of hard real-time systems such as avionics
and automobiles industry since the 80's~\cite{rp.twelve}.
%
The synchronous model trades power for reliability and has a simpler model
of time that suits most requirements of IoT applications.
%
On the one hand, this model cannot directly express time-consuming
computations, such as compression and cryptography algorithms, which are
typically either absent or delegated to auxiliary chips in the context of the
IoT.
%
On the other hand, all reactions to the external environment are guaranteed to
be computed in bounded time, ensuring that applications always reach an idle
state amenable to standby mode.
%
Overall, \CEU aims to offer a concurrent, safe, and expressive alternative to C
with the characteristics that follow:
%
\begin{description}
\item [Reactive:] code only executes in reactions to events and is idle most of
    the time.
\item [Structured:] programs use structured control mechanisms, such as
    \code{await} (to suspend a line of execution), and \code{par} (to combine
    multiple lines of execution).
\item [Synchronous:] reactions run atomically and to completion on each line of
    execution, i.e., there's no implicit preemption or real parallelism.
\end{description}
%
Structured reactive programming lets developers write code in direct style,
recovering from the inversion of control imposed by event-driven
execution~\cite{rp.deprecating,rp.rescala,sync_async.cooperative}.

@@

Note that the application of Figure~\ref{lst.inversion}.b relies on the driver
of Figure~\ref{lst.adc} to achieve transparent standby.
%
In \CEU, the burden to deal with standby is transferred to the device drivers,
which are write-once code written by specialists distributed with an SDK.
By transferring the work from the applications to the language level, novice
programmers never have to deal with standby explicitly.

In contrast, general-purpose languages typically provide low-power libraries to
deal with standby.
However, programmers still have to call these libraries explicitly,
characterizing a manual and error-prone mechanism.

Furthermore, we show in Figure~\ref{lst.inversion}.b that, when introducing
concurrency, the structure of the program remains sequential and amenable to
inference of the appropriate sleep mode.
In comparison with Arduino, whose main goal is to lower the entry barrier for
embedded development, \CEU also preserves the sequential structure for
concurrent applications.

>>>>

\CEU is a Esterel-based~\cite{ceu.tecs17} reactive programming language
targeting resource-constrained embedded systems~\cite{ceu.sensys13}.
%
It is grounded on the synchronous concurrency model, which has been
successfully adopted in the context of hard real-time systems such as avionics
and automobiles industry since the 80's~\cite{rp.twelve}.
%
The synchronous model trades power for reliability and has a simpler model
of time that suits most requirements of IoT applications.
%
On the one hand, this model cannot directly express time-consuming
computations, such as compression and cryptography algorithms, which are
typically either absent or delegated to auxiliary chips in the context of the
IoT.
%
On the other hand, all reactions to the external environment are guaranteed to
be computed in bounded time~\cite{ceu.sensys13}, ensuring that applications
always reach an idle state amenable to standby mode.
%
Overall, \CEU aims to offer a concurrent, safe, and expressive alternative to C
with the characteristics that follow:
%
\begin{description}
\item [Reactive:] code only executes in reactions to events and is idle most of
    the time.
\item [Structured:] programs use structured control mechanisms, such as
    \code{await} (to suspend a line of execution), and \code{par} (to combine
    multiple lines of execution).
\item [Synchronous:] reactions run atomically and to completion on each line of
    execution, i.e., there's no implicit preemption or real parallelism.
\end{description}
%
Structured reactive programming lets developers write code in direct style,
recovering from the inversion of control imposed by event-driven
execution~\cite{rp.deprecating,rp.rescala,sync_async.cooperative}.

@@

The application of Figure~\ref{lst.inversion}.b relies solely on the driver of
Figure~\ref{lst.adc} to achieve transparent standby.
%
In \CEU, the burden to deal with standby is transferred to the device drivers,
which are write-once code written by specialists distributed with an SDK.
By transferring the work from the applications to the language level, novice
programmers never have to deal with standby explicitly.
%
In contrast, general-purpose languages typically provide low-power libraries to
deal with standby.
However, programmers still have to call these libraries explicitly,
characterizing a manual and error-prone mechanism.

In Figure~\ref{lst.inversion}.b that, when introducing concurrency, the
structure of the program remains sequential and amenable to inference of the
appropriate sleep mode.
In comparison with Arduino, whose main goal is to lower the entry barrier for
embedded development, \CEU also preserves the sequential structure for
concurrent applications.

The synchronous model of \CEU provides logical parallelism for proper
separation of concerns, while avoiding the hassle of explicit synchronization
primitives (e.g., locks and mutexes).
%
Yet, asynchronous interrupts provide real-time responsiveness for
time-sensitive operations closer to the hardware.
